---
name: Jay Ballesteros-Cobián
date: "`r format(Sys.time(), '%B, %Y')`"
qualifications: MPP (candidate), Data Analyst, Strategy
position: Master of Public Policy (MPP) candidate at the University of Chicago
address: Chicago, IL
www: jballesterosc.github.io
email: jballesteros@uchicago.edu
github: jballesterosc
linkedin: jay-ballesteros
headcolor: "414141"
output: 
  vitae::awesomecv:
    keep_tex: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(vitae)
knitr::opts_chunk$set(echo = FALSE)
```

# About me

Master of Public Policy candidate at the University of Chicago with a quantitative focus on causal inference, program evaluation, and data analytics. Nine years of consulting and analytical experience supporting governments, multilateral-funded programs, and civil society organizations across Latin America, including projects for the U.S. Department of State, USAID-Chemonics, and AMEXCID. Proven ability to design monitoring and evaluation frameworks, conduct impact assessments of social programs, and translate evidence into policy recommendations. Proficient in Python, R, and SQL for end-to-end data solutions, from extraction and econometric modeling to visualization and decision-ready reporting.

# Education

```{r}
tribble(
~ degree, ~ uni, ~ loc, ~ dates, ~ details,

"Master of Public Policy (MPP), Candidate", "University of Chicago, Harris School of Public Policy", "Chicago, IL", "Sep. 2025 -- Present",
list(
"Quantitative concentration: Statistics and Econometrics. Microeconomics, and Analytical Politics",
"Upcoming coursework: Impact Evaluation, Programming for Data Analytics (Python), and Corporate Finance",
"Applied coursework using Python and R for data analysis, visualization, and experimental design"),

"B.A. International Relations", "Universidad de Colima", "Colima, México", "Aug. 2012 -- Jun. 2016", 
list("Emphasis on the Asia-Pacific region and international security",
     "Erasmus Mundus exchange at Universidad de León (Spain), coursework in finance and international trade"),
) %>%
detailed_entries(degree, dates, uni, loc, details)
```

# Experience

```{r}

tribble(
~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Senior Operations Analyst - Mature Cities", "DiDi Global Inc.", "Mexico City", "Apr. 2025 -- Nov. 2025",
    list("Led marketplace operations across three Mexican cities, overseeing a combined annual GMV of MXN 3.7B and a weekly incentives portfolio averaging MXN 4M",
         "Designed and executed randomized and quasi-experimental tests (A/B testing) to evaluate supply-demand balancing tactics on trips, fulfillment rates, user satisfaction, and other KPIs",
         "Built performance dashboards and automated reporting pipelines using Python, SQL, and Looker Studio; produced weekly analytical briefs for senior leadership with experiment results and action plans",
         "Partnered with regional and central teams (pricing, marketing, strategy, CX) to scale successful pilots and adapt evidence-based practices across markets"),

  "Co-founder and Data Analyst", "Datahué", "Mexico City", "Aug. 2023 -- Present",
    list(
    "Designed and implemented a MEAL strategy for Fundación Amparo, including 60+ performance indicators to track program outcomes and inform decision-making",
    "Consolidated 100+ databases into a centralized repository using Python, Google Apps Script, SQL, and Looker Studio; led a three-person team for data collection, cleaning, and evidence-based reporting to executive directors"),

  "Strategy Analyst", "Soy+ (soymas.la)", "Mexico City", "Mar. 2025 -- Apr. 2025",
    list(
    "Built automated dashboards connected via APIs to real-time databases; unified Mailchimp, Meta API, Supabase, and Vercel data into structured datasets for campaign evaluation and user segmentation"),

  "Senior Analyst", "C230 Consulting and Fundación IDEA", "Mexico City", "Jan. 2021 -- Mar. 2025",
    list("Delivered eight projects in security/justice reform, M\\&E, interinstitutional coordination, process optimization, market analysis and development for government counterparts and international organizations including the U.S. Department of State, USAID--Chemonics, and SANOFI",
    "Designed and executed an impact evaluation of the Impulso Nuevo León Loan Guarantee Program, analyzing changes in beneficiaries' sales and employment to assess program effectiveness on firm performance",
    "Created data-collection instruments for 80+ private-sector stakeholders in Honduras and El Salvador to inform the AMEXCID Jovenes Construyendo el Futuro program, supporting evidence generation for youth workforce policy",
    "Conducted a mixed-methods study on quality of life with Atopic Dermatitis for SANOFI, combining interviews, surveys, and quantitative analysis for evidence-based recommendations to national health institutions",
    "Developed institutional-strengthening toolkits for State Attorney General Offices; built measurement frameworks for recidivism monitoring in civic and criminal justice systems across multiple municipalities",
    "Advised on best practices in agriculture, youth workforce development, and migration across El Salvador, Honduras, Guatemala, and Mexico, identifying enabling conditions for replicable development initiatives"),

  "Founder and Data Analyst", "DataCol", "Colima, Mexico", "May 2020 -- Jul. 2021",
    list(
    "Led a government-funded analytics initiative for real-time COVID-19 monitoring, building time-series visualizations for the State Secretary of Health used in cabinet briefings; analyzed Google mobility datasets using Python to inform public health strategies"),

  "Consultant", "International Republican Institute", "Mexico City", "Aug. 2018 -- Jul. 2021",
    list(
    "Coordinated the Security with Justice Program in five Mexican states, engaging government counterparts, civil society, academia, and the private sector to promote ADR mechanisms; facilitated 30+ workshops training 1,100+ participants",
    "Managed MEL indicators in DevResults and Tableau; collected, analyzed, and reported open data on alternative justice across 15 states, identifying best practices for institutional strengthening"),

  "Competitiveness Manager", "Coparmex", "Colima, México", "Jun. 2016 -- Aug. 2018",
    list(
    "Directed capacity-building programs on business administration, productivity, and labor rights; produced research on security, justice, and economic trends for private-sector stakeholders")
) %>%
detailed_entries(role, dates, company, loc, details)
```

# Key skills and competencies

**Methods:** Causal inference and impact evaluation (diff-in-diff, RCTs, pre-post designs), experimental and quasi-experimental design (A/B testing), MEAL frameworks and program evaluation, quantitative and qualitative research (mixed methods)

**Tools:** Python, R, SQL, Excel, Looker Studio, Power BI, Tableau, DevResults; dashboarding, automation, and data engineering (APIs, pipelines); AI-assisted development tools (Claude Code, Copilot, Codex)

**Domain Expertise:** Productive development programs, innovation and entrepreneurship policy, security and justice reform, youth workforce development, public health; stakeholder management across Latin America

**Languages:** Spanish (native), English (professional fluency)
