---
name: Jay Ballesteros-Cobián
date: "`r format(Sys.time(), '%B, %Y')`"
qualifications: MPP (candidate), Data Analyst, Strategy
position: Senior Operations Analyst | Master of Public Policy candidate at the University of Chicago
address: Chicago, IL
email: jayballesterosc@gmail.com
linkedin: jay-ballesteros
github: jballesterosc
headcolor: "414141"gi
output: 
  vitae::awesomecv:
    keep_tex: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(vitae)
knitr::opts_chunk$set(echo = FALSE)
```

# About me

I am a quantitative analyst with a consulting background and eight years of experience turning complex data into high-impact decisions for governments, nonprofits, and tech firms. I combine structured problem-solving with in-depth technical skills in Python, R, SQL, and Looker Studio to develop end-to-end pipelines and interactive dashboards. I design and interpret A/B tests, difference-in-difference, and other quasi-experimental methods, distilling rigorous findings into actionable insights for marketplace optimization and evidence-based policy.

# Education

```{r}
tribble(
~ degree, ~ uni, ~ loc, ~ dates, ~ details,

"Master of Public Policy (MPP), Candidate", "University of Chicago, Harris School of Public Policy", "Chicago, IL", "Sep. 2025 -- Present",
list(
"Quantitative focus with statistics, microeconomics, analytical politics, program evaluation, and causal inference.",
"Applied coursework using Python and R for data analysis, visualization, and experimental design."),

"B.A. International Relations", "Universidad de Colima", "Colima, México", "Aug. 2012 -- Jun. 2016", "Emphasis on the Asia-Pacific region and international security.",

"Erasmus Mundus program", "Universidad de León", "León, Spain", "Sep. 2015 -- May. 2016", "Coursework focused on finance and international trade",
) %>%
detailed_entries(degree, dates, uni, loc, details)
```

# Experience

```{r}

tribble(
~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Senior Operations Analyst - Mature Cities", "DiDi Global Inc.", "Mexico City", "Apr. 2025 -- Present",
    list(
    "Led end-to-end marketplace operations across three key Mexican cities—overseeing a combined annual GMV of ~MXN 3.7 billion and managing a weekly incentives portfolio averaging MXN 4 million (3--10% of GMV)—optimizing trips, fulfillment rates, and user satisfaction through A/B tests and controlled experiments.",
    "Designed, executed, and tracked supply–demand balancing tactics (incentives, targeted communications, dynamic pricing) using Python, SQL, Looker Studio, and advanced spreadsheets.",
    "Analyzed business performance and local externalities, delivering weekly actionable insights to drive campaigns and growth.",
    "Partnered with regional and central teams (pricing, marketing, strategy, CX) to scale local pilots and adapt best practices, responding swiftly to operational disruptions."),
  
  "Co-founder and Data Analyst", "Datahué", "Mexico City", "Aug. 2023 -- Present",
    list(
    "Developed and implemented a Monitoring, Evaluation, Accountability, and Learning (MEAL) strategy for Fundación Amparo, designing 60+ performance indicators to track and evaluate impact.",
    "Consolidated 100+ databases into a centralized repository leveraging Python, Google Apps Script (Sheets), SQL, and Looker Studio, improving data integrity and operational efficiency.",
    "Led a three-person team for data collection, cleaning, and reporting; translated findings into actionable recommendations for program improvement."),

  "Strategy Analyst", "Soy+ (soymas.la)", "Mexico City", "Mar. 2025 -- Apr. 2025",
    list(
    "Aligned technology, commercial, and communications teams by translating requirements into actionable data-collection and analysis tools.",
    "Built automated Looker Studio dashboards connected via APIs to real-time databases to track conversion rates and weekly performance.",
    "Unified Mailchimp, Meta API, Supabase, and Vercel data into structured datasets for campaign evaluation, user segmentation, and behavior analysis."),

  "Senior Analyst", "C230 Consulting and Fundación IDEA", "Mexico City", "Jan. 2021 -- Mar. 2025",
    list("Delivered eight projects across security and justice reform, monitoring and evaluation, interinstitutional coordination, process optimization, and market analysis for clients including the U.S. Department of State, SANOFI, and USAID–Chemonics.",
    "Developed institutional--strengthening toolkits for State Attorney General Offices, leading process mapping and stakeholder engagement to improve investigative outcomes.",
    "Created data-collection instruments for more than 80 private--sector stakeholders in Honduras and El Salvador to inform the AMEXCID Jovenes Construyendo el Futuro project."),

  "Founder and Data Analyst", "DataCol", "Colima, Mexico", "May 2020 -- Jul. 2021",
    list(
    "Led a government-funded analytics initiative for real-time COVID-19 monitoring in Colima, building time-series visuals for infection and testing metrics.",
    "Analyzed Google mobility datasets to inform public health strategies; delivered Python tools to the State Secretary of Health for cabinet briefings.",
    "Assessed Facebook political advertising during the 2021 state elections, providing insights on reach, audience, and engagement."),

  "Consultant", "International Republican Institute", "Mexico City", "Aug. 2018 -- Jul. 2021",
    list(
    "Coordinated the Security with Justice Program in five states with government, civil society, academia, and the private sector to promote ADR mechanisms.",
    "Facilitated 30+ workshops on Restorative Justice and ADR, training 1,100+ participants at awareness and technical levels.",
    "Collected, analyzed, and reported open data on alternative justice across 15 states, identifying best practices and improvement areas."),

  "Competitiveness Manager", "Coparmex", "Colima, México", "Jun. 2016 -- Aug. 2018",
    list(
    "Directed capacity-building programs on business administration, productivity, tax reforms, and labor rights for affiliated companies.",
    "Produced research on security, justice, and economic trends to guide strategic decision-making for members and stakeholders.")
) %>%
detailed_entries(role, dates, company, loc, details)
```

# Key skills and competencies
  * Management consulting & strategy development
  * Data analytics & visualization (Python, R, SQL, Excel, Looker Studio, Power BI, Tableau)
  * Dashboarding, automation, and data engineering basics (APIs, pipelines)
  * Experimental and quasi-experimental methods (A/B testing, diff-in-diff)
  * Strategy and operations, KPI design, and reporting
  * MEAL frameworks and program evaluation
  * Quantitative & qualitative research (mixed methods)
  * Stakeholder management and cross-functional collaboration

